On multilevel constraint decomposition methods \\ for nonlinear variational inequalities

### 2021-07-08 paper/mcd2.tex
Stuck on coercivity of porous media form.

However, not all VIs correspond to optimization.  We give two such examples next, a linear advection-diffusion problem and then a nonlinear porous-medium-type problem, both of which are important in applications.  Each example is preceded by a lemma.

@@

\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive measure.  Let $\cV = H_0^1(\Omega)$ be the space of functions with value zero on $\partial_D\Omega$.  (For simplicity we consider zero boundary data.)  Choose a smooth and bounded velocity field $\bX$ on $\Omega$ such that $\Div \bX=0$, but additionally assume outward flow on the Neumann boundary, i.e.~$\bX \cdot \bn \ge 0$ at points of $\partial_N\Omega$.  For $u,v \in \cV = H_0^1(\Omega)$ and $\eps>0$ define
\begin{equation}
\ip{f(u)}{v} = \int_\Omega \eps \grad u \cdot \grad v - (\bX \cdot \grad u) v\,dx. \label{eq:advectiondiffusion}
\end{equation}
Consider VI \eqref{eq:vi} for any closed and convex $\cK \subset \cV$ and any $g\in\cV'$.  It is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous.  On the other hand, by Lemma \ref{lem:advectionskew}, which says that $s(u,v)$ is skew-symmetric up to a nonnegative addition, and by the outward flow assumption and the Poincar\'e inequality,
\begin{align*}
\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + s(u-v,u-v) \\
                    &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + \frac{1}{2} \int_{\partial_N\Omega} (u-v)^2 \bX\cdot\bn \ge \eps C \|u-v\|^2.
\end{align*}
Thus $f$ is 2-coercive, and so VI problem \eqref{eq:vi} is well-posed for any closed and convex $\cK \subset \cV$ and $g\in \cV'$.

IF $\bX=0$ THEN IT IS OPTIMIZATION BUT OTHERWISE NOT
\end{example}

In the above example we have assumed zero boundary data for simplicity.  Reference \cite{ChangNakshatrala2017} considers the above problem subject to a nonnegativity constraint ($\cK = \{v\ge 0\}$).

\begin{example}  \label{ex:porousgeneralization}  FIXME
\begin{equation}
\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u \cdot \grad v\,dx \label{eq:porousgeneralization}
\end{equation}
\end{example}

>>>>

However, not all VIs correspond to optimization.  We give two such examples next, first a linear advection-diffusion problem and then a nonlinear porous-medium-type problem.  Each of these examples, which are important in applications, is preceded by a lemma.

@@

\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive measure.  Let $\cV = H_0^1(\Omega)$ be the space of functions with value zero on $\partial_D\Omega$.  (For simplicity we consider zero boundary data.)  Choose a smooth and bounded velocity field $\bX$ on $\Omega$ such that $\Div \bX=0$, but additionally assume outward flow on the Neumann boundary, i.e.~$\bX \cdot \bn \ge 0$ at points of $\partial_N\Omega$.  For $u,v \in \cV = H_0^1(\Omega)$ and $\eps>0$ define
\begin{equation}
\ip{f(u)}{v} = \int_\Omega \eps \grad u \cdot \grad v - (\bX \cdot \grad u) v\,dx. \label{eq:advectiondiffusion}
\end{equation}
Consider VI \eqref{eq:vi} for any closed and convex $\cK \subset \cV$ and any $g\in\cV'$.  It is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous.  On the other hand, by Lemma \ref{lem:advectionskew}, which says that $s(u,v)$ is skew-symmetric up to a nonnegative addition, and by the outward flow assumption and the Poincar\'e inequality,
\begin{align*}
\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + s(u-v,u-v) \\
                    &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + \frac{1}{2} \int_{\partial_N\Omega} (u-v)^2 \bX\cdot\bn \ge \eps C \|u-v\|^2.
\end{align*}
Thus $f$ is 2-coercive, and so VI problem \eqref{eq:vi} is well-posed for any closed and convex $\cK \subset \cV$ and $g\in \cV'$.
\end{example}

However, if $\bX \ne 0$ then \eqref{eq:vi} for the functional \eqref{eq:advectiondiffusion} is not a (constrained) minimization problem.  Indeed, the Jacobian of $f$ is the bilinear form $J_f(u,v) = \int_\Omega \eps \grad u \cdot \grad v\,dx - s(u,v)$.  For nonzero advection this is not symmetric, thus it cannot be the Hessian of an objective $F$.  (For $u,v$ which are zero on all of $\partial \Omega$, $J_f(u,v) - J_f(v,u) = -2 s(u,v)$, and we may then construct $u,v$ so this result is nonzero anywhere $\bX$ is nonzero.)

References \cite{Bueler2021conservation,ChangNakshatrala2017} consider such advection-diffusion VI problems over $\cK = \{v\ge 0\}$.

\begin{example}  \label{ex:porousgeneralization}  FIXME. I THINK THIS IS NOT GOING TO WORK.  COERCIVITY IS SUPER FRAGILE W.R.T. THIS CHANGE?
\begin{equation}
\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u \cdot \grad v\,dx \label{eq:porousgeneralization}
\end{equation}
\end{example}
